### 2020-03-09 document/manuscript.tex
Travis build: 67

@@

@@

What is the take-home message from this article? What are the
recommendations?

>>>>

\section{Towards Open and Reproducible Genomic Research: Lessons from
OpenScienceKE}\label{towards-open-and-reproducible-genomic-research-lessons-from-openscienceke}

\subsection*{Abstract}\label{abstract}
\addcontentsline{toc}{subsection}{Abstract}

Open, collaborative and reproducible research -- Open Science -- has a
great potential for advancing science. However, the training in our
local universities does not equip students with the tools to practice
open science. However, to work in the open and collaborate, your
collaborators should be equipped to use the tools that you use. The main
barrier to working open, therefore, is the lack of awareness of the
collaboration tools and the skills required to utilize these tools.
Therefore, to fill the gap through an open science community, funded by
a Mozilla Mini-grant -- OpenScienceKE, we are promoting open science
among bioinformatics students and researchers in the Nairobi area by
training using this model: sensitize, train, hack and collaborate. This
model first sensitizes on open science practices through seminars,
trains on open science tools through workshops, facilitates hands-on
application of the tools through hackathons, and finally fosters a
community of open science enthusiasts through meetups.

OpenScienceKE sought to address the following problems: the lack of
awareness of open science practices and tools within the Bioinformatics
community in Kenya; the poor adoption of open science practices in
Bioinformatics; and the absence of research to establish the state of
affairs in adopting open science in Kenya. From the OpenScienceKE
hackathon, we managed to create an open resource that the students could
use to figure out where they can cost-effectively publish open access.
In addition, through literature search and data mining, we observed a
growing interest in open science practices in Kenya but the lack of
awareness and skills hinder the adoption. The use of preprints for
research dissemination haven't caught up in Kenya; out of the 20,069
papers downloaded from bioRXiv, only 18 have Kenyan authors, a majority
of which are as a result of international (16) collaborations. We also
observed a lack of incentives and policy in academic and research
institutions to support open science. The fear of being scooped and the
competitive spirit within the scientific community are also major
barriers to working in the open.

The first iteration of the model which focused on academic institutions
set the foundation for next phase: promote the open and reproducible
science in research institutions. This model provides the framework for
the adoption of open science practices within the institution and others
in the future. As genomic research data generated in Africa grows, there
is a need for the adoption of open science practices in data storage,
reproducible pipelines and collaborative research. We propose this
approach, which develops the necessary infrastructure within research
institutions, and builds human capacity through the model: sensitize,
train, hack and collaborate. Promotion of open science in Africa
recognizes the future direction of research and OpenScienceKE is growing
the culture and practice in the research active region.

@@

Our Core message is to present a clear outlook on the adoption of Open
Science practice by African-based scientists, starting with Kenya.

@@

What is the take-home message from this article? What are the
recommendations? - The need for a framework to guide the adoption of
open science practices. Take note of the bariers and provide
recommendations. - The need for low-cost publishing - The need for
policies on open science that can be implemented by various
insituitions. For example, we can provide a template that can be adopted
by most instituitions
